---
title: "Assignment 4"
author: "Bhakti Sangoi"
date: "February 8, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

#GitHub
- https://github.com/sangoibhakti/NEU-DA5020.git

#Reading Libraries
```{r}
library(tidyverse)
library(readxl)
library(stringr)
library(forcats)
library(ggthemes)
```

#Reading Datasets
```{r}
farmers_market <-read_csv("farmers_market.csv") #Reading Farmers market dataset
kyfp <- read_excel("kyfprojects.xls") #Reading Know your farmers and Know your food project
```

#Warm-up
This dataset stores city and state in different columns, what if you want to
print out city and state in the format "City, State"?
```{r}
CityState_frmt = str_c(farmers_market$city, ', ',farmers_market$State) #Combine City,State
```

#Question 1:
Cleanup the Facebook and Twitter column to let them contain only the facebook username or twitter handle name.
- replace "https://www.facebook.com/pages/Cameron-Park-Farmers-Market/97634216535?ref=hl" with "Cameron-Park-Farmers-Market", 
-"@21acres" with "21acres

```{r}
fm_fb <- select(farmers_market, Facebook) 
fm_vec <- as_vector( fm_fb) #Creating a vector
df1 <- data_frame(facebook = character())
# Clean up for Facebook
    for(x in 1:NROW(fm_vec))  #forming For loop to check each pattern with per data
    {
      p= fm_vec[[x]]
      a1 = 'https:[/][/]www.facebook.com[/]pages[/]'
      b1 = 'https:[/][/]www.facebook.com[/]'
      c1 = 'www.facebook.com[/]'
      d1 = 'facebook.com[/]'
      e1 = 'facebook[/]'
      f1 = 'www.facebook.com[/]pages[/]'
      g1 = 'https:[/][/]www.'
      h1 = 'www.'
          ss1 =if(grepl(a1,p)==TRUE){gsub(a1, "", p)}
          else if(grepl(b1,p)==TRUE){gsub(b1, "", p)}
          else if(grepl(c1,p)==TRUE){gsub(c1, "", p)}
          else if(grepl(d1,p)==TRUE){gsub(d1, "", p)}
          else if(grepl(e1,p)==TRUE){gsub(e1, "", p)}
          else if(grepl(f1,p)==TRUE){gsub(f1, "", p)}
          else if(grepl(g1,p)==TRUE){gsub(g1, "", p)}
          else if(grepl(h1,p)==TRUE){gsub(h1, "", p)}
          else {p}
      df1[x, ]=ss1
    }
Facebook_username <- data_frame(gsub("/.*","",as.character(df1$facebook)))
Facebook_username #displaying facebook username
farmers_market$Facebook <- Facebook_username

#Clean up for Twitter
fm_twit <- select(farmers_market, Twitter)
fmt_vec <- as_vector(fm_twit) #Creating a vector
df2 <- data_frame(twitter = character())
    for(x in 1:NROW(fmt_vec)) #forming For loop to check each pattern with per data
    {
        q= fmt_vec[[x]]
        m1 = 'https:[/][/]twitter.com[/]'
        n1 = 'www.twitter.com'
        l1 = '@'
            tt1 =if(grepl(m1,q)==TRUE){gsub(m1, "", q)}
            else if(grepl(n1,q)==TRUE){gsub(n1, "", q)}
            else if(grepl(l1,q)==TRUE){gsub(l1, "", q)}
            else {q}
        df2[x, ] = tt1
    }
df2 #displaying the twitter username
farmers_market$Twitter <- df2


```

#Question 2:
Clean up the `city` and `street` column. Remove state and county names from the `city` column and consolidate address spellings to be more consistent (e.g. "St.", "ST.", "Street" all become "St"; "and" changes to "&", etc...).
```{r}
#Cleaning Street
farmers_market <- farmers_market %>% 
  mutate(
          street = street %>% 
            str_replace_all(c( "\\b(and)\\b" = "&", #Replace 'and' by '&'
                               "\\bAve?(\\.|\\b)" = "Avenue", #Replacing 'Ave' by 'Avenue'
                               "\\bBlvd(\\.|\\b)" = "Boulevard" #Replacing 'Blvd' by       'Boulevard'
                          ))
        )

farmers_market$street <-gsub("(St|st|Sts.|Street|Streets|street|streets)","Street",farmers_market$street) #Replacing all matches of street like Street/st/sts./ by "Street"
farmers_market$street

#Cleaning City
farmers_market$city <- gsub(",\\s([A-z]|[a-z]){2}","",farmers_market$city)
farmers_market$city
```

#Question 3:
Create a new data frame (tibble) that explains the online presence of each state's farmers market. I.e., how many percentages of them have a facebook account? A twitter account? Or either of the accounts? (Hint: use the `is.na()` function)
```{r}
farmers_percentage<- farmers_market %>% 
                     group_by(State) %>% #Group by State 
                    #Summarize columns to find percentage of accounts by State
                     summarise(FaceBook = ((sum(!is.na(Facebook)))/n())*100, 
                     Twitter = ((sum(!is.na(Twitter)))/n())*100,
                     Youtube= ((sum(!is.na(Youtube)))/n())*100,
                     Website = ((sum(!is.na(Website)))/n())*100,
                     Other = ((sum(!is.na(OtherMedia)))/n())*100)
farmers_percentage
```

#Question 4:
Some of the farmer market names are quite long. Can you make them shorter by using the `forcats::fct_recode` function? Create a plot that demonstrates the number of farmers markets per location type. The locations should be ordered in descending order where the top of the graph will have the one with the highest number of markets.
```{r}
farmers_market<-farmers_market %>% 
                        as_tibble()%>%
                        mutate(Abbr_Location =  #Using mutate to create the new column and                                                  keeping the old location as it is
                                 recode(Location,
  "Faith-based institution (e.g., church, mosque, synagogue, temple)"= "Religional Inst.",
  "Closed-off public street" = "Public St.",
  "Private business parking lot" = "Private Parking",
  "Local government building grounds" = "Local Govnt.",
  "Federal/State government building grounds" = "Federal_Govnt Bld.",
  "Educational institution" = "Education Inst.",
  "Healthcare Institution" = "Health Inst.",
  "On a farm from: a barn, a greenhouse, a tent, a stand, etc" = "On_Farm",
  "Co-located with wholesale market facility"= "Wholesale Market" 
    )) 

farmers_market$Abbr_Location #A new column with Abbreviated location is created.

Farmersmarket_location <- farmers_market %>% 
                          group_by(Abbr_Location) %>% #Group by Abbreviated Location 
                          count(Abbr_Location) %>%
                          filter(!is.na(Abbr_Location)) # Remove NA's

#Demonstarting Number of Farmers Market by Abbreviated Location
ggplot(data=Farmersmarket_location, mapping=aes(y=reorder(Abbr_Location,n),x=n)) +
geom_point()
```

#Question 5:
Write code to sanity check the `kyfprojects` data. For example, does `Program Abbreviation` always match `Program Name` for all the rows? (Try thinking of your own rules, too.)
```{r}

#Does Program Abbreviation always match Program Name
unique(kyfp$`Program Name`) #21 Entries
unique(kyfp$`Program Abbreviation`) #19 Entries

#The unique no. of Program Name does not match with Program Abbreviation, hence it fails the sanity test.

#Kyfp dataset consists of 2379 values

#Generating additional rules

#Year has only digits
year_digits <- grepl("^[[:digit:]]*$",kyfp$Year)
count_dig <- sum(year_digits < 7) #Gives the count for values satisfying above rule.
count_dig
#It satisfies the rule

#Zipcode is only numeric
zip1<-grepl("^[0-9]+$",kyfp$Zip)
count_zip1<-sum(zip1<7) #Gives the count for values satisfying above rule
count_zip1
#It satisfies the rule

#Zipcode has only 5 digits
zip2<-grepl("\\d{5,}",kyfp$Zip)
count_zip2<-sum(zip2<7) #Gives the count for values satisfying above rule
count_zip2
#It satisfies the rule

#Check if USDA Agency has characters
usda_char <- grepl("[A-Z]", kyfp$`USDA Agency`)
count_usda <- sum(usda_char < 7) #Gives the count for values satisfying above rule
count_usda
#It satisfies the rule

#State has only 2 characters
state2<-nchar(kyfp$State)
count_state2<-sum(state2<7) #Gives the count for values satisfying above rule
count_state2
#It satisfies the rule

#Recipient has all alphabets
rec <- grepl("[A-Z]", kyfp$Recipient)
count_rec <-sum(rec < 7) #Gives the count for values satisfying above rule
count_rec
#It satisfies the rule

#Funding type has alphabets
fund_type <-grepl("[A-Z]",kyfp$`Funding Type`)
count_ft <-sum(fund_type < 7) #Gives the count for values satisfying above rule
count_ft
#It satisfies the rule
```