---
title: "Assignment 5"
author: "Bhakti Sangoi"
date: "February 10, 2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

#GitHub
- https://github.com/sangoibhakti/NEU-DA5020.git

#Reading Libraries
```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(lubridate)
```

#Reading Farmers Market and Kyfprojects Dataset
```{r}
farmers_market <-read_csv("farmers_market.csv") #Reading Farmers market dataset
kyfp<- read_excel("kyfprojects.xls") #Reading Kyfproject dataset
```

#Question 1:
Add a new column `Season1Days` that contains the number of days a market is opened per week (for the dates it is open).
```{r}
#used str_count function: to count the number of days market is open and create a new column per season to store the result
farmers_market$Season1Days<-str_count(farmers_market$Season1Time,";")
farmers_market$Season1Days
farmers_market$Season2Days<-str_count(farmers_market$Season2Time,";")
farmers_market$Season2Days
farmers_market$Season3Days<-str_count(farmers_market$Season3Time,";")
farmers_market$Season3Days
farmers_market$Season4Days<-str_count(farmers_market$Season4Time,";")
farmers_market$Season4Days
```

#Question 2:
Add a new column `WeekendOpen` indicating whether a market opens during weekends in `Season1.
```{r}
#To find if market is open on weekend by using grepl 
farmers_market$WeekendOpen<-grepl("Sat|Sun",farmers_market$Season1Time)
#WeekendOpen counts
str_count(farmers_market$WeekendOpen,"TRUE")
Weekend_Open_TotalCount <- sum(str_count(farmers_market$WeekendOpen,"TRUE"))
Weekend_Open_TotalCount
#3191 farmers market are open in Season 1.
```

#Question 3:
Find out which markets close before 6PM, and which open only for fewer than 4 hours a day. For simplicity, consider only `Season1Time`. For markets with different open hours across a week, use the average length of open hours for the days they actually open.
```{r}

#Selecting Market Name AND Season 1 from dataset
n <- farmers_market %>% select(MarketName,Season1Time) 
p1 <-str_replace_all(n$Season1Time,"^[A-Za-z:]+[ ]|^[A-Za-z]+[:]","")
q1 <- str_replace(p1,";","") #clean-up data and fetching time
#View(as.data.frame(q1))
#extracting start time
r1 <- str_extract(p1,"^[0-9]+[:][0-9]+[ ][A-Za-z]+") 
#View(as.data.frame(r1))
parse_date_time(r1, "%I:%M %p") #Setting it in a date-time format using parse

#extrating end time
s1 <- str_extract(q1 , "[0-9]+[:][0-9]+[ ][A-Za-z]+$")
#View(as.data.frame(s1)) 
parse_date_time(s1, "%I:%M %p")  #Setting it in a date-time format using parse
n$Season1Time <- q1

#Constaint that market closes before 6.00PM
constraint <- parse_time("6:00 PM")

#Adding Start and End time column
start_endtime<-mutate(n , begin = r1, end = s1)

#Displaying Season1Time, StartTime and EndTime
final_markettime <- filter(start_endtime, abs(parse_time(begin)-parse_time(end)) < 14400, parse_time(end) < constraint)
final_markettime
```

#Question 4:
The seasons are not standardized and would make analysis difficult. Create four new columns for four seasons (Spring, Summer, Fall, Winter), indicating whether a market is available in that season. Also, create two additional columns `HalfYear` and `YearRound` to identify those who open across seasons. Define "half year" and "year round" on your own terms, but explain them before you write the code (or as comments in your code). (Hint: you may want to create even more auxiliary columns, `Season1BeginDate` and `Season1EndDate` for example.)

```{r}
#Defining:
##Year_Round: Farmers market is open throughout the year considering for Season 1. 
##Half_Year: Farmers market is open from start date of Season 1 for 6 months span.

#Clean-up data
farmers_market$Season1Date <- str_replace_all(farmers_market$Season1Date, c("January"="01/01/2015","February"="02/01/2015","March"="03/01/2015","April"="04/01/2015",   "May"="05/01/2015","July"="07/01/2015", "June"="06/01/2015","August"="08/01/2015",         "September"="09/01/2015","October"="10/01/2015","November"="11/01/2015","December"="12/01/2015"))

#Cleaning Season 1 Start Date
Season1_Start <- str_extract(farmers_market$Season1Date,"^[0-9]+[/][0-9]+[/][0-9]+")

#Cleaning Season 1 End Date
Season1_End <- str_extract(farmers_market$Season1Date,"[0-9]+[/][0-9]+[/][0-9]+$")

#Replacing Date with year
s1 <- str_replace(Season1_Start,"[0-9]+$","2017")
s1 <- parse_date(s1, "%m/%d/%Y")
e1 <- str_replace(Season1_End,"[0-9]+$","2017")
e1 <- parse_date(e1 , "%m/%d/%Y")

# create date ranges for the four seasons
Winter_Start <- as.Date("2017-12-01")
Spring_Start <- as.Date("2017-03-01")
Summer_Start <- as.Date("2017-06-01")
Fall_Start   <- as.Date("2017-09-01")
#Intervals
winter_season <- interval(Winter_Start, Spring_Start + years(1))
spring_season <- interval(Spring_Start,Summer_Start)
summer_season <- interval(Summer_Start,Fall_Start)
fall_season   <- interval(Fall_Start,Winter_Start)
Half_Year <- interval(s1, s1+months(6) )
Year_Round <- interval(Winter_Start, Fall_Start+months(3) )
winter_season
spring_season
summer_season
fall_season

farmers_market<- farmers_market %>% mutate(
      
                Season1_startdate = s1,
                Season1_enddate = e1,
                # consider all dates to be of the same year, along with end of the year    situation
                Season1_enddate = if_else(Season1_enddate < Season1_startdate,
                                Season1_enddate + years(1),
                                Season1_enddate),
                Season1_span = interval(Season1_startdate, Season1_enddate),
                #returning True/False using int_overlaps
                winter_season = int_overlaps(Season1_span, winter_season),
                spring_season = int_overlaps(Season1_span,spring_season),
                summer_season = int_overlaps(Season1_span,summer_season),
                fall_season = int_overlaps(Season1_span, fall_season),
                Year_Round = int_overlaps(Season1_span,Year_Round),
                Half_Year = int_overlaps(Season1_span,Half_Year)
                )

View(farmers_market)
```

#Question 5:
_Open question_: explore the new variables you just created. Aggregate them at different geographic levels, or some other categorical variable. What can you discover?
```{r}
#using coord_flip: Flipping cartesian coordinates X and Y axis along with geom_bar to make height of the bar proportional to no. of cases in each group.


##A)Farmers Market is open maximum number of days in the state of California followed by the state of New York
ggplot(data = farmers_market)+
geom_bar(mapping = aes(x = State, fill = Season1Days))+
coord_flip()

##B)In Fall, maximum markets are open in different states
ggplot(data = farmers_market)+
geom_bar(mapping = aes(x = State, fill = fall_season))+
coord_flip()

##C)In Spring, maximum markets are open in different states
ggplot(data = farmers_market)+
  geom_bar(mapping = aes(x = State, fill = spring_season))+
  coord_flip()

##D)In Summer, maximum markets are open in different states
ggplot(data = farmers_market)+
geom_bar(mapping = aes(x = State,fill = summer_season))+
coord_flip()

##E)In Winter, maximum markets are closed in different states
ggplot(data = farmers_market)+
geom_bar(mapping = aes(x = State, fill = winter_season))+
coord_flip()

##F)Different States have farmers market open and closed based on seasons based on Year round analysis
ggplot(data = farmers_market)+
  geom_bar(mapping = aes(x = State, fill = Year_Round))+
  coord_flip()

##G)For half year markets are open most of the time in a particular time frame.
ggplot(data = farmers_market)+
  geom_bar(mapping = aes(x = State, fill = Half_Year))+
  coord_flip()

##H)Statistical Analysis is done on which state has the high and low number of farmers market Open on Weekends.
ggplot(data = farmers_market)+
geom_bar(mapping = aes(x = State, fill = WeekendOpen))+
coord_flip()
```